Project Setup

### 2024-09-13 archive.qmd
Rework setup, README, and archival sections
Following Malika's suggestions

---
title: "Archive"
lang: "en"
bibliography: literature.bib
---

@@

Additional guidance is provided by @Daquin2023 and @Garijo2022. We will cover the archival using the Software Heritage archive and Zenodo.

>>>>

---
title: "Archive Your Project"
lang: "en"
bibliography: literature.bib
---

## Where to Deposit Software

@@

Additional guidance is provided by @Daquin2023 and @Garijo2022.

## Using Zenodo to Archive Your Project

In the following we will demonstrate how to use Zenodo to archive one's project and obtain a DOI.

::: {#cau-sandbox .callout-caution}
### Zenodo Sandbox

Note that during this tutorial, we will be using Zenodo's sandbox instance located at [`sandbox.zenodo.org`](https://sandbox.zenodo.org/) in order to learn how to use it while not wasting Zenodo's resources. For real projects, use the production instance at [`zenodo.org`](https://zenodo.org/), which requires a separate account.
:::

First, create an account on the Zenodo Sandbox by clicking on _Sign up_ in the upper right corner and submitting the provided form. Then, confirm your email address (also check your spam folder if you cannot find the confirmation email).

Now you can follow the guide "[Create new upload](https://help.zenodo.org/docs/deposit/create-new-upload/)", taking into account the following notes specific to each step:

- __Add/remove files__: To circumvent the file limit on Zenodo, you can bundle your project folder in one ZIP file and upload that instead.
- __Fill in metadata__
  - __Digital Object Identifier__ (DOI): Usually, you will want a DOI, therefore you need to choose _No_ for the question whether you already have a DOI for this upload, and click on _Get a DOI now!_.
  - __Resource type__: Select the most appropriate one, for example, _Publication / Preprint_.
  - __Licenses__: If your project only contains content by you and if you previously only chose one license, you can select that by choosing _Add standard_. Otherwise, click on _Add custom_ and describe how each file is licensed. If you have been using REUSE to record license information, you can run enter `SPDX Document` as title and run the reuse tool with `reuse spdx -o reuse.spdx` to create the file `reuse.spdx`. Then you can paste the content of `reuse.spdx` into the description field of the custom license as it contains information about the license of each file.

If you previously uploaded your project to a public Git repository (which we didn't cover during this tutorial), you could also add the new DOI to your README, for example, under the section "Citation".

## Final Remarks

There are situations in which you are not allowed to share your data publicly. In this case, consider using a repository specifically for research data, which later grants access for valid uses individually.

### 2024-09-13 make_readme.qmd
Rework setup, README, and archival sections
Following Malika's suggestions

## Creating a README

@@

Badges and project status

@@

:   What steps need to be taken to run the project? What software needs to be installed? Mention all dependencies here that are not explicitly managed by `renv`, such as the system dependencies of R packages as well as the version of Quarto.

    We also know that we need Quarto to create the PDF and that we installed the `apaquarto` extension, so let's find out these versions as well:^[Luckily, the extensions are included in the project folder, so technically their version is already recorded in the project's files.]

    ```{.bash filename="Terminal"}
    quarto --version
    quarto list extensions
    ```

    ```txt
    [✓] Checking tools....................OK
          TinyTeX: (not installed)
          Chromium: (not installed)

    [✓] Checking LaTeX....................OK
          Using: Installation From Path
          Path: /usr/local/texlive/2024/bin/x86_64-linux
          Version: 2024
    ```

    We put all the information obtained about system dependencies and their versions in the README. Sharing the install scripts that were provided by `pak` is an additional courtesy to your readers.

@@

Create your README now as the file `README.md`.

>>>>

## What does a README contain?

@@

Badges/Project Status

@@

:   What steps need to be taken to run the project? What software needs to be installed? Mention all dependencies here that are not explicitly managed by `renv`, such as the system dependencies of R packages as well as the version of Quarto. An R package's system dependency is any additional software that you need to install on your computer in order to use a particular R package. For example, the R package [`reticulate`](https://rstudio.github.io/reticulate/) allows to run Python code from within R. However, in order to actually use it, one has to additionally install Python itself as it does not come together with `reticulate` -- rather, it is a system dependency. See @sec-dependencies for additional information.

@@

## Installation/Dependencies {#sec-dependencies}

We also know that we need Quarto to create the PDF, so let's find out its version as well:

```{.bash filename="Terminal"}
quarto --version
```

If you installed `apaquarto` or any other Quarto extension, one can query their versions as follows:^[Luckily, the extensions are included in the project folder, so technically their version is already recorded in the project's files.]

```{.bash filename="Terminal"}
quarto list extensions
```

```txt
[✓] Checking tools....................OK
      TinyTeX: v2024.09
      Chromium: (not installed)

[✓] Checking LaTeX....................OK
      Using: TinyTex
      Path: /home/r155953/.TinyTeX/bin/x86_64-linux
      Version: 2024
```

## Create It!

Create your README now as the file `README.md`. If you feel stuck, you can have a look at the following examples:

::: {#tip-name .callout-tip collapse="true"}
### Name

```{.md .code-overflow-wrap filename="README.md"}
# Penguin Paper

```
:::

::: {#tip-project-status .callout-tip collapse="true"}
### Badges/Project Status

```{.md .code-overflow-wrap filename="README.md"}
[![Project Status: Unsupported – The project has reached a stable, usable state but the author(s) have ceased all work on it. A new maintainer may be desired.](https://www.repostatus.org/badges/latest/unsupported.svg)](https://www.repostatus.org/#unsupported)
```
:::

::: {#tip-description .callout-tip collapse="true"}
### Description

```{.md .code-overflow-wrap filename="README.md"}
This project contains the Quarto manuscript of our study on penguins. It is written in R and uses `renv` to track its dependencies.
```
:::

::: {#tip-installation .callout-tip collapse="true"}
### Installation/Dependencies

``````{.md .code-overflow-wrap filename="README.md"}
## Dependencies

This manuscript requires the following system software to be installed. In addition, we provide the version numbers this manuscript has last been run with:

- [Quarto](https://quarto.org/docs/download/) 1.6.9
- GNU Make 4.4.1
- Pandoc 3.3
- TinyTeX 2024.09
- [R](https://cloud.r-project.org/) 4.4.1

On Fedora Linux, Make and Pandoc can be installed as follows:

```bash
dnf install -y make pandoc
```

Quarto and R can be installed using the links provided. TinyTeX can be installed using Quarto by entering the following into the terminal:

```bash
quarto install tinytex
```

All R packages that this project requires are managed using [`renv`](https://cran.r-project.org/package=renv). Therefore, `renv` needs to be installed first, by entering the following in the R console:

```r
install.packages("renv")
```

Next, one can open a new R session in the root directory of this project and run the following, which should install all required R packages at their recorded versions:

```r
renv::restore()
```
``````
:::

::: {#tip-usage .callout-tip collapse="true"}
### Usage

`````{.md .code-overflow-wrap filename="README.md"}
## Usage

The most important file in this project is `Manuscript.qmd` which contains the text of the article as well as the code for its computations. It is accompanied by the following files:

- `Bibliography.bib`: bibliographic references used in the manuscript
- `Data.csv`: a data set containing the simplified `palmerpenguins` data created using `write.csv(palmerpenguins::penguins, "dat.csv", row.names = FALSE)`
- `data_dictionary.html`: a dictionary to the data file, created using `create_data_dictionary.R`

The folder `_extensions` contains the `apaquarto` extension which is used to typeset the PDF accoording to APA guidelines.

The manuscript can be rendered to PDF using the following command:

```bash
quarto render Manuscript.qmd
```
``````
:::

::: {#tip-citation .callout-tip collapse="true"}
### Citation

```{.md .code-overflow-wrap filename="README.md"}
## Citation

Please cite this draft as follows:

> Zerna & Scheffel (2024): "A Study on Penguins: A Minimal Reproducible Example". Unpublished manuscript.
```
:::

::: {#tip-citation .callout-tip collapse="true"}
### License

```{.md .code-overflow-wrap filename="README.md"}
## License

The manuscript file and the bibliograhpic references are made available by us under [CC BY 4.0](https://creativecommons.org/licenses/by/4.0/). The data set by Kristen Gorman has been published under [CC0 1.0](https://creativecommons.org/publicdomain/zero/1.0/). `apaquarto` has also been published under [CC0 1.0](https://creativecommons.org/publicdomain/zero/1.0/). All other files in this project are available under [CC0 1.0](https://creativecommons.org/publicdomain/zero/1.0/).
```
:::

### 2024-09-13 setup.qmd
Rework setup, README, and archival sections
Following Malika's suggestions

The manuscript also uses the Quarto extension "[apaquarto](https://wjschne.github.io/apaquarto/)", which typesets documents according to the requirements of the @APA2020. It can be installed in the project using the following command:

>>>>

The manuscript also uses the Quarto extension "[apaquarto](https://wjschne.github.io/apaquarto/)", which typesets documents according to the requirements of the American Psychological Association [-@APA2020]. It can be installed in the project using the following command:
